\Large \bf \Toolname: Precise False Sharing Detection for COTS Binaries

### 2014-11-04 HotOS'15/paper.tex
Second commit. Done most of the related work, although I am missing some things that would probably be useful things to have, but that I don't diectly draw inspiration from (e.g. Pluto, Callgrind, Simics). Fleshed out the intro a bit. Now onto design section.

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf \Toolname: Precise Detection of False Sharing in COTS Binaries}
\numberofauthors{3}
\author{
% 1st. author
\alignauthor
Peter Goodman\\
       \email{pag@cs.toronto.edu}
\and
% 2nd. author
\alignauthor
Ashvin Goel\\
       \email{asvhin@eecg.toronto.edu}
% 3rd. author
\alignauthor
Angela Demke Brown\\
       \email{demke@cs.toronto.edu}
}
\maketitle
\begin{abstract}

Hello
\end{abstract}

@@

In this work, we present \Toolname, a system for precisely detecting the sources of false sharing in
multi-threaded, commercial off-the-shelf (COTS) binaries. This is important because false sharing
can limit the scalability of multi-threaded applications on cache-coherent multiprocessors
\cite{ImpactOfFalseSharing}. The key insight of our approach is that it begins by focusing on what
data is being accessed, and then uses that information to progressively narrow down on what code
is participating in a data race.

@@

\section{Related Work}\label{sec:background}
\Toolname{} draws on techniques from existing race detectors and false sharing detectors. This
is appropriate because some forms of false-positives for race detectors are caused by false
sharing. That is, where race detectors focus on two threads concurrently accessing the same
data, false sharing detectors focus on two threads concurrently accessing different data, but
on the same cache line.

@@

\Toolname{} is a dynamic, precise, variable-granularity, mixed-mode false-sharing detector. It draws inspiration
from a number of existing race detectors, including LiteRace \cite{LiteRace}, MultiRace \cite{MultiRace},
RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and DataCollider \cite{DataCollider}.

@@

MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
lockset-based approach. It operates at an object granularity, and therefore reports false-positives. It is unique
in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
access pattern of a thread (no-acces, read-only, read-write). \Toolname{} uses a similar pointer swizzing approach, but
in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
include a type identifier that is unique to a $\langle PC,size \rangle$ pair. This improves the diagnosis of false-sharing,
because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
two threads accessing two nearby objects of the same (or different) types.

@@

\section{Design}\label{sec:design}

@@

>>>>

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf \Toolname: Precise False Sharing Detection for COTS Binaries}
\numberofauthors{3}
\author{
% 1st. author
\alignauthor
Peter Goodman\\
       \email{pag@cs.toronto.edu}
\and
% 2nd. author
\alignauthor
Ashvin Goel\\
       \email{asvhin@eecg.toronto.edu}
% 3rd. author
\alignauthor
Angela Demke Brown\\
       \email{demke@cs.toronto.edu}
}
\maketitle

%\begin{abstract}
%\end{abstract}

@@

% Key ideas:
%		1) 	Shadow memory as a proxy for memory accesses, not as a meta-data store.
%		2) 	Tagged reads as a way to avoid cache contention on shadow memory, but still
%			highlight the essential memory access behaviour.
%		3)	Proportional heap sampling, by using type-specific information to decide where
%			to place hardware watchpoints.

False-sharing is a performance-degrading problem that limits the scalability of multithreaded applications
\cite{ImpactOfFalseSharing}. This problem is increasingly important in the multi-core era. The level of
concurrency in applications is growing to match the additional processing power provided by increasing
core counts on modern processors. Despite this rise in importance, detecting and diagnosing false
sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
sharing is often indistinguishable from true sharing (including race conditions). Also, diagnosing the cause
of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.

In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
in multi-threaded, commercial off-the-shelf (COTS) binaries. The key insights of our approach is that
it iteratively discovers and specializes instrumentation to detect false sharing at increasingly finer
granularities. Our work addresses three limitations of prior work:
\begin{itemize}
\item They introduce extra cache coherency traffic, which limits scalability and hides performance
problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.

\item They are unable to selectively detect false sharing on only specific types of objects or specific
code.

\item They are imprecise: they cannot distinguish between false sharing, true sharing (including
data races), and and pseudo sharing. This can result in high false-positive rates that decrease the
usefulness of the tools.

\end{itemize}

\paragraph{Contributions}
This paper makes the following contributions:
\begin{itemize}
\item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
false sharing  in multi-threaded, CTOS binaries. We believe that operating directly on binaries is
important because it allows one to verify observed performance problems using the same binaries
that are running in a production environment.

\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
efficient determination of memory access behavior using shadow memory. Unlike existing shadow
memory systems, proxy memory is not treated as a mutable meta-data store. This is important
for false sharing detection as it avoids the extra serialization and coherence traffic that is introduced
by dynamically manipulating shadow state.

\item It describes a two new, data-centric sampling approaches: \begin{inparaenum}[i)]
\item uniform heap sampling, and
\item selective heap sampling.
\end{inparaenum} Uniform heap sampling enables \Toolname{} to sample objects in the heap uniformly
according to their type. This is beneficial because it improves coverage: if 90\% of false sharing occurs because
of accesses to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the
other 90\% of data. Once data of interesting has been narrowed down, \Toolname{} can apply selective heap
sampling, which focuses sampling on specific types of objects or data accessed by specific code (or both).
\end{itemize}

@@

% --------------------------------------------------------------------------------------

\section{Related Work}\label{sec:background}
\Toolname{} primarily draws on techniques from existing race detectors and false sharing detectors.
This is appropriate because some forms of false-positives for race detectors are caused by false
sharing. That is, where race detectors focus on two threads concurrently accessing the same
data, false sharing detectors focus on two threads concurrently accessing different data, but
on the same cache line.

@@

Under this categorization scheme, \Toolname{} is a dynamic, precise, variable-granularity, mixed-mode
false-sharing detector. It draws inspiration from a number of existing race detectors, including LiteRace
\cite{LiteRace}, MultiRace \cite{MultiRace}, RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and
DataCollider \cite{DataCollider}.

@@

MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
lockset-based approach. It operates at an object granularity, and therefore reports false-positives. It is unique
in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
access pattern of a thread (read-write > read-only > no-access). \Toolname{} uses a similar pointer swizzing approach, but
in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
include a type identifier that is unique to a $\langle PC,size \rangle$ pair. This improves the diagnosis of false-sharing,
because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
two threads accessing two nearby objects of the same (or different) types.

@@

Similar to race detectors, false-sharing detectors can be sampling-based or comprehensive. Comprehensive approaches
can be categorized  into simulation-based and instrumentation-based approaches. Simulation-based approaches focus on
simulating the full cache hierarchy and can distinguish fine-grained events like false sharing due to set-associativity and
false sharing due to different data being accessed on the same cache line. Instrumentation-based approaches focus more
on mechanisms for estimating or detecting cache line contention.

\paragraph{Sampling-based}
Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
counters and monitoring events. It represents the state-of-the-art in performance analysis tools and is able to provide detailed,
architecture-specific performance-tuning suggestions. Unlike \Toolname{}, VTune is unable to analyze only a portion of the code,
or only the portion of code that accesses specific shared data structures.

\paragraph{Comprehensive}

% Simulation based approaches:
% TODO: Callgrind/Cachegrind?
% TODO: Simics?
% TODO: Pluto?

\newcommand{\DrContention}{Zhao \emph{et al.}}

\DrContention{} \cite{DrContention} implement a dynamic false and true sharing detector using the DynamoRIO \cite{DynamoRIO}
DBT system and Umbra shadow memory framework \cite{Umbra}. Unlike \Toolname{}, \DrContention{} use fixed-granularity
shadow memory to track cache line ownership by processor cores. The drawback of this approach is that it necessitates checking
and updating the shadow memory on every memory access, which introduces its own form of cache line contention. \Toolname{}
avoids this issue by using shadow memory and tagged reads as a proxy for memory access behavior instead of as a meta-data store.

Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Like \Toolname{}, it takes a pipeline-based
approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
their repair mechanism, which transparently re-maps memory accesses at a byte granularity.

SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.

% --------------------------------------------------------------------------------------
\section{Design}\label{sec:design}

% --------------------------------------------------------------------------------------
\section{Implementation}\label{sec:implementation}
\Toolname{} is implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
\cite{AddressWatchpoints} framework.

@@

% Answer the following questions:
%		1)	What are DataReactor's overheads over time?
%		2) 	How sensitive is DataReactor to different sampling rates?s
%		3)	How effective is DataReactor as discovering false sharing?
%		4)	Are the cases of detected false sharing high impact?
